---
title: "pfars2 Package Vignette"
output: rmarkdown::html_vignette
author: Ibrahim Badalov
vignette: >
  %\VignetteIndexEntry{Package_Vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(pfars2)
```

## General Information About pfars2 Package.
This package is intented to read, manipulate and visualize datasets from US National
Highway Traffic Safety Administration's Fatality Analysis Reporting System.
There are five functions:
1. make_filename()
2. fars_read()
3. fars_read_years()
4. fars_summarize_years()
5. fars_map_state()


## Definitions and Examples

#### make_filename()
The main purpose of this function is to create names for datasets to read.
It takes year argument to create filename for relevant year. It should be
pointed out that file name has a default structure so 
you can only manipulate the year.
```{r}
make_filename(year = 2013)
```


#### fars_read()
This function is to read dataset from working directory. A dataset can be
read in two ways, writing dataset name manually and using make_filename function.
While giving both examples only run one of them.
```{r}
#fars_read(filename = "accident_2013.csv")
fars_read(filename = make_filename(year = 2013))
```


#### fars_read_years()
This function has similiar functionality with fars_read(), however it has a series of
factors. This is able to read a couple of datasets at one call storing in a list.
It takes a vector consists of years rather than dataset names and only reads year
and month columns from datasets. It can be merged doing some manipulation 
as given in the example.
```{r warning=FALSE}
fars_read_years(c(2013,2014))
suppressPackageStartupMessages(library(dplyr))
fars_read_years(years = c(2013,2014)) %>%
  bind_rows() %>% head(10)
```


#### fars_summarize_years()
This function of the package pfars2 helps to read and summarize counts by year
giving years as columns. As argument it takes a numeric vector of years.
The example below explains far better purpose of the function.
```{r}
fars_summarize_years(years = c(2013,2014))
```


#### fars_map_state()
This function has a bit more different task comparing others. It allows user to visualize
state map with data points. It takes state number and years as arguments.
Example below illustrates in better way.
```{r}
library(maps)
fars_map_state(state.num = 50, year = 2013)
```
